---
title: "Action Item responses"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=TRUE, warning=FALSE, paged.print=FALSE, fig.width = 7)
options(encoding = "UTF-8")

##Important!! Enter credential information. Include last "/" of the baseurl
##ENTER BASE URL, USERNAME, PASSWORD
baseurl<-"https://hrhr.dhis2.org/dhis/"
username<-"brian"
password<-"eReg!234"


##Load required packages
required_packages<-c("dplyr","stringr","lubridate","httr","assertthat","readr","stats","DT","ggplot2","utils","datasets","methods","base","knitr","reshape2")      

is_installed <- function(mypkg) is.element(mypkg, installed.packages()[,1])

load_or_install<-function(required_packages) {
  for(package_name in required_packages)  {
    if(!is_installed(package_name))     {
       install.packages(package_name)  }
    library(package_name,character.only=TRUE,quietly=TRUE,verbose=FALSE)
  }
}

load_or_install(required_packages)
  
  
##download libraries and test login
require(readr)
require(assertthat)
require(httr)
loginDHIS2<-function(baseurl,username,password) {
  url<-paste0(baseurl,"api/me")
  r<-GET(url,authenticate(username,password))
  assert_that(r$status_code == 200L) }
if(loginDHIS2(baseurl,username,password)){
  print("successfully logged in")
}else{
  stop("could not log in! Please check url, username and password")
}

```

Analysis date `r Sys.Date()` for `r baseurl`

...Pulling the datastore items and responses ...


```{r schedule, echo=FALSE, warning=FALSE, include=FALSE}
options(encoding = "UTF-8")

##pull interaction keys
require(httr)
require(jsonlite)
dataKey_api<-"api/dataStore/actionFeedback/"
url<-paste0(baseurl,dataKey_api)

keys<-as.data.frame(fromJSON(content(GET(url), "text"), flatten = TRUE), stringsAsFactors=FALSE)
colnames(keys)<-"id"


#for each program ID, pull scheduled events and add to a dataframe
#first, create an empty dataframe
interactions<-data.frame()
e<-c("positive","negative","positiveFeedbackMessages","negativeFeedbackMessage","id")
for (k in e) interactions[[k]]<-as.character()

url<-paste0(baseurl, "api/dataStore/actionFeedback/",keys$id[1])

for (value in 1:length(keys$id)){

url<-paste0(baseurl, "api/dataStore/actionFeedback/",keys$id[value])
action<-fromJSON(content(GET(url), "text"))

positive<-action$positive
negative<-action$negative

if(is.null(unlist(action$positiveFeedbackMessages))){
    posfeedback<-""}else{
      posfeedback<-action$positiveFeedbackMessages
    }

if(is.null(unlist(action$negativeFeedbackMessages))){
    negfeedback<-""}else{
      negfeedback<-action$negativeFeedbackMessages
    }
negfeedback
posfeedback

action1<-as.data.frame(cbind(positive,negative,posfeedback,negfeedback,"id"=keys$id[value]),stringsAsFactors=FALSE)
names(action1)<-e



if (length(action1)>0){
  interactions<-rbind(interactions,action1)
}
next()

}
interactions$id<-as.integer(interactions$id)

interactions



```



```{r Pull all validation results, echo=FALSE, warning=FALSE, include=FALSE}

#name API
vresults_api<-"api/validationResults.json?paging=false&fields=id,period,organisationUnit,created,validationRule"
url<-paste0(baseurl,vresults_api)


#pull validation results as data frame
vresults<-as.data.frame(fromJSON(content(GET(url), "text"), flatten = TRUE), stringsAsFactors=FALSE)

#rename output headers
x<-sub('validationResults.','', names(vresults))
x<-sub('.id$','',x)
names(vresults)<-x


```

```{r Merge context data, echo=FALSE, warning=FALSE, include=FALSE}

Sys.setlocale(category = "LC_ALL", locale = "Arabic")

#now pull org units and validationRule IDs

ou_api<-"api/organisationUnits.json?paging=false&fields=id,level,name,parent"
url<-paste0(baseurl,ou_api)

orgunits<-fromJSON(content(GET(url), "text"), flatten = TRUE)[[1]]

library(dplyr)
vresults_2<-dplyr::left_join(vresults, orgunits, by=c("organisationUnit"="id"))

parents<-orgunits %>%
  select(id, name) %>%
  rename(parent.name = name, parent.id = id)

grandparents<-orgunits %>%
  select(name, id) %>%
  rename(grandparentname = name) %>%
  mutate_at("id",as.character)


vresults_2<-left_join(vresults_2, parents, by="parent.id")
vresults_2<-left_join(vresults_2, grandparents, by=c("parent.id"="id"))


## now pulling valRules and instructions
val_api<-"api/validationRules.json?paging=false&fields=id,code,name,description,instruction"
url<-paste0(baseurl,val_api)
valRules<-fromJSON(content(GET(url), "text"), flatten = TRUE)[[1]]

valRules<- valRules %>%
  rename(valRuleName = name, valRuleCode = code)


vresults_2<-left_join(vresults_2, valRules, by=c("validationRule"="id"))


```


```{r lastly, merge the results with contextual data, echo=FALSE}

outputs<-left_join(vresults_2, interactions, by=c("id"))
write_excel_csv(outputs, "all_actionItem_responses.csv", append=FALSE)

library(dplyr)
outputs_sum<- outputs %>%
  mutate_at(c("negative", "positive"), as.integer) %>%
  na.omit() %>%
  group_by(validationRule, valRuleName, instruction) %>%
  summarise(positive = sum(positive), negative = sum(negative)) %>%
  arrange(desc(positive, negative)) %>%
  select(validationRule, valRuleName, instruction, positive, negative)


outputs_responses <- outputs %>%
  group_by(name, valRuleName, instruction) %>%
  na.omit() %>%
  arrange(desc(positiveFeedbackMessages, negativeFeedbackMessage, valRuleName)) %>%
  select(name, valRuleName, instruction, positiveFeedbackMessages, negativeFeedbackMessage)

outputs_byOU <- outputs %>%
  mutate_at(c("negative", "positive"), as.integer) %>%
  group_by(grandparentname, parent.name,name) %>%
  na.omit() %>%
  summarise(positive = sum(positive), 
            negative = sum(negative)) %>%
  select(grandparentname, parent.name,name, positive, negative)

library(dplyr)
outputs_forchart <- outputs %>%
  mutate_at(c("negative", "positive"), as.integer) %>%
  na.omit() %>%
  group_by(period) %>%
  summarise(positive = sum(positive), 
            negative = sum(negative)) %>%
  select(period, negative, positive)

outputs_forchart$period<-as.Date(paste0(as.character(outputs_forchart$period), '01'), format='%Y%m%d')


```
Pulled `r length(outputs)` responses to action items, available in CSV file of working directory.

Now we output the results in tables...

####Here is a summary of responses to action items

```{r Datatable outputs1, include=TRUE}
DT::datatable(outputs_sum, 
              escape = FALSE,
              extensions = 'Buttons', 
              options = list(
                pageLength = 5,
                dom = 'Blfrtip', 
                buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))


```

####Responses to feedback messages by org unit
```{r Datatable outputs2, include=TRUE}

DT::datatable(outputs_byOU,
              escape = FALSE,
              extensions = 'Buttons', 
              options = list(
                pageLength = 5,
                dom = 'Blfrtip', 
                buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))


```

####Clinics written response to feedback messages
```{r Datatable outputs3, include=TRUE}

DT::datatable(outputs_responses,
              escape = FALSE,
              extensions = 'Buttons', 
              options = list(
                pageLength = 5,
                dom = 'Blfrtip', 
                buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))


```


####Chart of responses over time
```{r over time}

library(ggplot2)
library(reshape2)

df2<-melt(outputs_forchart, id=c("period"))


ggplot(df2, aes(x=period, y=value, group=variable)) +
        geom_col(position="dodge", aes(fill=variable)) +
        labs(title="action item feedback delivered by month", ylab="responses") +
        scale_x_date(date_labels = "%Y%m", date_breaks = "1 month") +
           geom_text(
               aes(label = value, y = value + 0.05),
                   vjust = 0
           )


```